---
title: "PART 2 - Baseline OLS Model"
format: html
---

## Setup

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(modelsummary)
```

## Load data

```{r}
data <- read_excel("../data/jtrain3.xlsx") %>%
  mutate(
    re78 = as.numeric(re78),
    re75 = as.numeric(re75),
    re74 = as.numeric(re74),
    train = as.numeric(train),
    educ = as.numeric(educ),
    age = as.numeric(age),
    black = as.numeric(black),
    hisp = as.numeric(hisp),
    trre75 = as.numeric(trre75)
  )

glimpse(data)
```

## Baseline OLS regression

```{r}
ols1 <- lm(
  re78 ~ train + re75 + educ + age,
  data = data
)

summary(ols1)
```

## Regression table

```{r}
modelsummary(ols1)
```

# PART 3 - Extended Model with Demographic Controls

## Extended OLS model

```{r}
ols2 <- lm(
  re78 ~ train + re75 + re74 + educ + age + black + hisp,
  data = data
)

summary(ols2)
```

## Regression table

```{r}
modelsummary(list(
  "Baseline model" = ols1,
  "Extended model" = ols2
))
```

# PART 4 - Interaction Effects

## Model with interaction between training and previous earnings

```{r}
ols3 <- lm(
  re78 ~ train + re75 + trre75 + educ + age,
  data = data
)

summary(ols3)
```

## Model comparison

```{r}
modelsummary(
  list(
    "Baseline" = ols1,
    "Extended" = ols2,
    "Interaction" = ols3
  ),
  stars = TRUE
)
```

## Model fit comparison

```{r}
data.frame(
  Model = c("Baseline", "Extended", "Interaction"),
  R_squared = c(
    summary(ols1)$r.squared,
    summary(ols2)$r.squared,
    summary(ols3)$r.squared
  ),
  Adjusted_R_squared = c(
    summary(ols1)$adj.r.squared,
    summary(ols2)$adj.r.squared,
    summary(ols3)$adj.r.squared
  )
)
```

# Part 5 - Model Diagnostics

## Packages

```{r}
library(car)
library(lmtest)
library(sandwich)
library(corrplot)
```
```{r}
corr_data <- data %>%
  select(re75, re74, educ, age)

corr_matrix <- cor(corr_data, use = "complete.obs")

corr_matrix
```

```{r}
corrplot(corr_matrix, method = "color")
```

## a) Variance Inflation Factor (VIF)

```{r}
vif(ols2)
```

## b) Breusch-Pagan test

```{r}
bptest(ols2)
```

## c) Robust standard errors

```{r}
coeftest(
  ols2,
  vcov = vcovHC(ols2, type = "HC1")
)
```

## d) RESET test

```{r}
resettest(ols2)
```

# Part 6 - Earnings Changes and Causal Interpretation

## a) Construct earnings change

```{r}
data <- data %>%
  mutate(
    delta_re = re78 - re75
  )
```
  
## Average earnings changes by training status

```{r}
data %>%
  group_by(train) %>%
  summarise(
    mean_delta = mean(delta_re, na.rm = TRUE),
    sd_delta = sd(delta_re, na.rm = TRUE),
    median_delta = median(delta_re, na.rm = TRUE),
    n = n()
  )
```
  
## b) Regression using earnings changes

```{r}
ols_delta <- lm(
  delta_re ~ train + educ + age,
  data = data
)

summary(ols_delta)
```

# Part 7 - Heterogeneous Treatment Effects
## a) Estimate preferred model separately by race

```{r}
data_black <- data %>%
  filter(black == 1)

data_nonblack <- data %>%
  filter(black == 0)

ols_black <- lm(
  re78 ~ train + re75 + re74 + educ + age + hisp,
  data = data_black
)

ols_nonblack <- lm(
  re78 ~ train + re75 + re74 + educ + age + hisp,
  data = data_nonblack
)

summary(ols_black)
summary(ols_nonblack)
```
